ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-02-26 pyked.tex
edits and additions to intro

%====================================================================
\section{Introduction}
%
Fundamental combustion experiments provide vital data for understanding
fuel chemistry and validating chemical kinetic models. Important measured quantities
include autoignition delay times, laminar flame speeds, and species profiles.
However, despite both the importance and abundance of such information in the literature,
the combustion\slash chemical kinetics community lacks an accepted, commonly used
standard for recording and sharing data from fundamental combustion experiments.

Interestingly, this contrasts the situation for chemical kinetic models, where the
CHEMKIN~\autocite{Kee:1996ck} format dominates. Competing standards such as the
Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code}
lag behind considerably, although use of the former continually grows due to Cantera's
open availability.
\todo[inline, author=Bryan]{This reads more as a footnote, maybe consider putting it there instead?}
% Similarly, % with more space, discuss JANAF/NASA polynomial standard ?

Instead, such data is typically found in under-documented
comma-separated value files and Excel spreadsheets, or contained in PDF tables
rather than as supplementary material associated with a paper.
In the worst case, data is only available in figures and must be digitized by the use of software such as WebPlotDigitizer.
\todo{give a citation here}
% I think we're already stirring the pot enough, no need to call out
% specific people
% \todo{call out specific examples of these...?}{}
These practices limit the wide use of the valuable, extensive experimental data
available in the literature. Frenklach~\autocite{Frenklach:2007bm} further explained the
benefits of standardized, widely available data for combustion.
\todo[author=Bryan,inline]{It would maybe be good to give a bit of Frenklach's perspective here, rather than just throwing in the citation}

While the community has not widely adopted a standard format or the practice of openly \todo{this second bit feels like a slightly separate point, and it makes the second clause here a bit awkward}
sharing data, some researchers and groups attempted to solve the first problem.
Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
\textbf{M}od\textbf{e}l) data format~\autocite{Frenklach:2007bm,You:2011hy} and an associated
online database, the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
combustion experimental data, along with kinetic models and calculated quantities,
using the Extensible Markup Language (XML) standard.

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represented an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
prevented wide adoption: the PrIMe standard does not require or support all information
needed to simulate an experiment, including a machine-readable definition of ignition or a
standard way to express the rapid compression machine (RCM) compression process; and
the XML format, intended more as a markup language than data format, presents a barrier
to creating and working with files.
\todo{Also because the standard was closed?}

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
an XML-based format that builds on PrIMe. While ReSpecTh adds important features, it still
suffers from the same usability issues as PrIMe.
\todo{slightly more detail?}

In this work, we present a new, open-source, human- and machine-readable data standard for fundamental combustion
experiments, \ck{}, and offer tools for easily working with data encoded in this format.
Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
and in this work we further formalize and develop the standard. We also discuss a related
Python software tool for validating and working with \ck{} files: \pk{}.
Our motivations are similar those of the PrIMe and ReSpecTh teams, although
less ambitious in scope\todo{its not clear to me how this is less ambitious than the other projects}: the \ck{} project will enable easy sharing and use of fundamental combustion data, for the (primary) purposes of
developing and validating predictive chemical kinetic models.
Furthermore, we plan to share all data and software openly, and cultivate a community of
user-developers.

\todo[inline, author=Bryan]{We should probably also mention the CloudFlame project from KAUST.}

@@

The \yaml|value| can be a string, integer, or floating point number. The type is
inferred from the \yaml|key| when the value is read. The \ck{} format is
designed to include all of the information necessary to simulate a given
experiment and can be broken into two main sections. The first section details
all of the ``meta'' information about the experiment and the \ck{} file. In this
section, the common \yaml|author| mapping is used for any fields where an author
must be specified. The \yaml|author| mapping contains the following fields:
%
\begin{itemize}
    \item \yaml|name| (required, string): The author's full name
    \item \yaml|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
    of the \ck{} file
    \item \yaml|file-version| (required, integer): The version of the \ck{} file
    \item \yaml|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yaml|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
        mappings
        \item \yaml|journal| (required, string): The name of the publishing
        journal
        \item \yaml|year| (required, integer): The year of publication
    \end{itemize}
    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
    delay| or \yaml|flame speed|
    \item \yaml|apparatus| (required, string): The type of apparatus used to
    perform the experiment
\end{itemize}

@@

The \ck{} file that will be used in this example can be found in the
\python|tests| directory of the PyKED repository,
\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: (1, 0)
chemked-version: 0.0.1
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the data point is from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, there is just a single datapoint in this file, which codes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure, compression time, ignition type, and the volume history that specifies
how the volume of the reactor varies with time, simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
  - temperature: 297.4 kelvin
    ignition-delay: 1.0 ms
    pressure: 958.0 torr
    composition:
      - species-name: H2
        InChI: 1S/H2/h1H
        mole-fraction: 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        mole-fraction: 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        mole-fraction: 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        mole-fraction: 0.63125
    ignition-type:
      target: pressure
      type: d/dt max
    compression-time: 38.0 ms
    volume-history:
      time:
        units: s
        column: 0
      volume:
        units: cm3
        column: 1
      values:
        - [0.00E+000, 5.47669375000E+002]
        - [1.00E-003, 5.46608789894E+002]
        - [2.00E-003, 5.43427034574E+002]
        - [3.00E-003, 5.38124109043E+002]
        ...
\end{yamlbox}
%
The values for the \yaml|volume-history| are truncated to save space. A user
might like to load the information from this file to perform a simulation to
calculate the ignition delay using Cantera. All of the information required to
perform this simulation is present in the \ck{} file, with the exception of a
chemical kinetic mechanism for H$_2$/CO combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

# Load the ChemKED file and retrieve the first element of the
# datapoints list, which is an instance of the DataPoint class
ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]
\end{pythonbox}
%
The initial temperature, pressure, and mixture composition can be read from the
instance of the \python|DataPoint| class. \pk{} uses Pint \python|Quantities| to
store values with units, while Cantera expects a floating point value in SI
units as input. Therefore, we use the built-in capabilities of Pint to convert
the units from those specified in the \ck{} file to SI units:
%
\begin{pythonbox}
T_0 = dp.temperature.to('K').magnitude
P_0 = dp.pressure.to('Pa').magnitude
X_0 = dp.get_cantera_mole_fraction()

# Load the mechanism and set the initial state of the mixture
gas = ct.Solution('h2-co-mechanism.cti')
gas.TPX = T_0, P_0, X_0

# Create the reactor and the outside environment
reac = ct.IdealGasReactor(gas)
env = ct.Reservoir(ct.Solution('air.xml'))
\end{pythonbox}
%
To apply the effect of the volume trace to the \python|IdealGasReactor|, a
\python|Wall| is installed between the reactor and the environment and assigned
a velocity. Although we do not show the details here, a reference implementation
of a class that computes a wall velocity given the volume history of the reactor
is available in CanSen~\autocite{cansen}, in the
\python|cansen.profile.VolumeProfile| class.
%
\begin{pythonbox}
time = dp.volume_history.time
volume = dp.volume_history.volume
ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))

netw = ct.ReactorNet([reac])

# Integrate for 50 ms
while netw.time < 0.05:
    netw.step()
\end{pythonbox}
%
Although not shown in this example, the user would typically store information
about the \python|IdealGasReactor| (e.g., temperature, pressure, mass fractions)
at each time step of the integration for post-processing.

@@

The \ck{} file used in this example can be found online.\todo{add url}{} The
data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
In this \ck{} file, multiple data points are specified with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yaml|common-properties| section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    - species-name: t-butanol
      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
      mole-fraction: 0.003333333
    - species-name: O2
      InChI:  1S/O2/c1-2
      mole-fraction: 0.04
    - species-name: Ar
      InChI:  1S/Ar
      mole-fraction: 0.956666667
  ignition-type:  &ign
    target: OH* concentration
    type: 1/2 max
\end{yamlbox}
%
This block utilizes the ability of YAML files to define an anchor using the
\yaml|&| symbol and refer to that section later using the \yaml|*| symbol, as
can be seen in the definition of the \yaml|datapoints| section:
%
\begin{yamlbox}
datapoints:
  - temperature: 1459 kelvin
    ignition-delay: 347 us
    pressure: 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1389 kelvin
    ignition-delay: 756 us
    pressure: 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1497 kelvin
    ignition-delay: 212 us
    pressure: 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1562 kelvin
    ignition-delay: 105 us
    pressure: 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}

In this example, we would like to run constant volume simulations at each of the
pressure and temperature conditions in the \yaml|datapoints| list. Once again,
the \ck{} file specifies all the information required for the simulations except
for the chemical kinetic mechanism.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

@@

# Since the composition is specified by the common-properties,
# just take the composition of the first DataPoint
X_0 = ck.datapoints[0].get_cantera_mole_fraction()

# Loop through each of the DataPoints
for dp in ck.datapoints:
    T_0 = dp.temperature.to('K').magnitude
    P_0 = dp.pressure.to('Pa').magnitude
    gas.TPX = T_0, P_0, X_0
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    # Define ignition delay as T_0 + 400 K
    while reac.T < T_0 + 400:
        netw.step()

    print('The ignition delay for T_0={}, P_0={} is: {} seconds'.format(
        T_0, P_0, netw.time
    ))
\end{pythonbox}
%
Note that the ignition delay in this example is defined as equal to $T_0 +
400\,\mathrm{K}$, a simplified definition used for this example. In general, the
user could post-process the concentration information from the simulation to
determine one-half the maximum OH* concentration to match the experimental
definition of ignition delay.
%====================================================================
\section{Conclusions and future work}
%
\todo[inline]{both to do}

>>>>

%====================================================================
\section{Introduction}
%
Fundamental combustion experiments provide vital data for understanding
fuel chemistry and validating chemical kinetic models. Important measured quantities
include autoignition delay times, laminar flame speeds, and species profiles.
However, despite both the importance and abundance of such information in the literature,
the combustion\slash chemical kinetics community lacks an accepted, commonly used
standard for recording and sharing data from fundamental combustion experiments.\footnote{
Interestingly, this contrasts the situation for chemical kinetic models, where the
CHEMKIN~\autocite{Kee:1996ck} format dominates. Competing standards such as the
Cantera~\autocite{Cantera:2.3.0} CTI format or FlameMaster~\autocite{FlameMaster:ref,FlameMaster:code}
lag behind considerably, although use of the former continually grows due to Cantera's
open availability.
% Similarly, % with more space, discuss JANAF/NASA polynomial standard ?
}

Instead, such data is typically found in under-documented
comma-separated value files and Excel spreadsheets, or contained in PDF tables
rather than as supplementary material associated with a paper.
In the worst case, data is only available in figures and must be digitized by the
use of software such as WebPlotDigitizer.
\todo{give a citation here}
% I think we're already stirring the pot enough, no need to call out
% specific people
% \todo{call out specific examples of these...?}{}
These practices limit the wide use of the valuable, extensive experimental data
available in the literature. Frenklach~\autocite{Frenklach:2007bm} further explained the
benefits of standardized, widely available data for combustion.
In brief, much of combustion\slash chemical kinetics research drives towards the ultimate goal
of developing predictive kinetic models---and ultimately this depends on a community
infrastructure for data and methods.

The combustion community has not yet widely adopted a standard format for experimental data,
but some researchers and groups attempted to solve this problem.\todo[author=Kyle]{this still doesn't read quite right}{}
Most notably, Frenklach developed the PrIMe (\textbf{Pr}ocess \textbf{I}nformatics
\textbf{M}od\textbf{e}l) data format~\autocite{Frenklach:2007bm,You:2011hy} and an associated
online database, the PrIMe Data Warehouse~\autocite{PrIMe}. PrIMe files encode fundamental
combustion experimental data, along with kinetic models and calculated quantities,
using the Extensible Markup Language (XML) standard.

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} represents an admirable \todo{Maybe its just me, but this feels a bit sarcastic} first attempt at standardization, but a number of limitations
prevent wide adoption: the PrIMe standard does not require or support all information
needed to simulate an experiment, including a machine-readable definition of ignition or a
standard way to express the rapid compression machine (RCM) compression process; and
the XML format, intended more as a markup language than data format, presents a barrier
to creating and working with files.
Furthermore, the closed and opaque nature of the PrIMe standard and associated Data Warehouse
limit contributions and community development of these and compatible tools.

More recently, Varga et al.~\autocite{Varga2015a,Varga2015b} developed the ReSpecTh standard,
an XML-based format that builds on PrIMe. ReSpecTh adds important features that make files
better standalone representations of experimental data---i.e., more informative by themselves
rather than in concert with a larger system.
For example, where PrIMe used internal bibliographic references, ReSpecTh added a field for
typical bibliographic data. The ReSpecTh standard also provides machine-readable formats
for better specifying ignition experiments, including a definition of autoignition and
either volume histories for RCMs or pressure-rise in shock tubes.
However, being XML-based, ReSpecTh experimental files still suffer from the same usability
issues as PrIMe, and the standard remains closed.
\todo[author=Kyle,inline]{how's this?}

Olm and colleagues used their ReSpecTh-based database to quantify the performance of literature
hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models, and more recently
develop improved models for hydrogen\slash syngas~\autocite{Varga:2016gj} and
ethanol~\autocite{Olm:2016et}. In all four cases, they converted numerous experimental datasets
from the literature into the ReSpecTh standard.

Building on the lack of an accepted data standard, limited examples exist of open sharing
of such data. The Clean Combustion Research Center at King Abdullah University of Science
and Technology (KAUST) started the CloudFlame project in 2013~\autocite{Goteng:2013cf}\todo{a few more refs needed}{}
(\url{https://cloudflame.kaust.edu.sa}). CloudFlame does serve as an openly accessible database
for experimental data, available in a standard CSV format, but also goes beyond this by
providing a cloud infrastructure for running simulations based on stored models and data.
While an admirable (and useful) effort, CloudFlame represents another closed system controlled
by a specific institution rather than the community.

In this work, we present a new, open-source, human- and machine-readable data standard for fundamental combustion
experiments, \ck{}, and offer tools for easily working with data encoded in this format.
Niemeyer recently introduced an initial version of the \ck{} format~\autocite{Niemeyer:2016wf},
and in this work we further formalize and develop the standard. We also discuss a related
Python software tool for validating and working with \ck{} files: \pk{}.
Our motivations are similar those of the PrIMe and ReSpecTh teams:
the \ck{} project will enable easy sharing and use of fundamental combustion data,
for the (primary) purposes of developing and validating predictive chemical kinetic models.
However, we make usability a major design focus, and plan to share all data and software openly
to cultivate a community of user-contributors.

@@

The \yaml|value| can be a string, integer, or floating point number. The type is
inferred from the \yaml|key| when the value is read. The \ck{} format is
designed to include all of the information necessary to simulate a given
experiment and can be broken into two main sections. The first section details
all of the ``meta'' information about the experiment and the \ck{} file. In this
section, the common \yaml|author| mapping is used for any fields where an author
must be specified. The \yaml|author| mapping contains the following fields:
%
\begin{itemize}
    \item \yaml|name| (required, string): The author's full name
    \item \yaml|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
\todo[author=Kyle,inline]{one difference ReSpecTh called out from PrIMe is that
the latter added a copyright line in all files, claiming copyright on the data...
do we want to be incendiary and point out that you can't copyright data?}
%
\begin{itemize}
    \item \yaml|file-author| (required, \yaml|author|-type mapping): The author
    of the \ck{} file
    \item \yaml|file-version| (required, integer): The version of the \ck{} file
    \item \yaml|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yaml|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yaml|authors| (required, sequence): A sequence of \yaml|author|
        mappings
        \item \yaml|journal| (required, string): The name of the publishing
        journal
        \item \yaml|year| (required, integer): The year of publication
    \end{itemize}
    \item \yaml|experiment-type| (required, string): One of \yaml|ignition
    delay| or \yaml|flame speed|
    \item \yaml|apparatus| (required, string): The type of apparatus used to
    perform the experiment
\end{itemize}

@@

The \ck{} file that will be used in this example can be found in the
\python|tests| directory of the PyKED repository,
\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
Examining that file, we find the first section specifies the information about
the \ck{} file itself
%
\begin{minted}{yaml}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: (1, 0)
chemked-version: 0.0.1
fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
\end{minted}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the data point is from the work of \textcite{Mittal2006a}:
%
\begin{minted}{yaml}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
  fake-key: 0 # Again here because Atom's syntax highlighting is busted :(
\end{minted}

Finally, there is just a single datapoint in this file, which codes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure,
%
\begin{minted}{yaml}
datapoints:
  - temperature: 297.4 kelvin
    ignition-delay: 1.0 ms
    pressure: 958.0 torr
    composition:
      - species-name: H2
        InChI: 1S/H2/h1H
        mole-fraction: 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        mole-fraction: 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        mole-fraction: 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        mole-fraction: 0.63125
    ignition-type:
      target: pressure
      type: d/dt max
    compression-time: 38.0 ms
    volume-history:
      time:
        units: s
        column: 0
      volume:
        units: cm3
        column: 1
      values:
        - [0.00E+000, 5.47669375000E+002]
        - [1.00E-003, 5.46608789894E+002]
        - [2.00E-003, 5.43427034574E+002]
        - [3.00E-003, 5.38124109043E+002]
        - 0 # Dammit Atom
        ...
\end{minted}

The values for the \yaml|volume-history| are truncated to save space. The user
can load this file into \pk{} and run a simulation in Cantera with the following
code:

\begin{minted}{python}
import cantera as ct
from pyked import ChemKED

ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]

T_0 = dp.temperature.to('K').magnitude
P_0 = dp.pressure.to('Pa').magnitude
X_0 = dp.get_cantera_mole_fraction()

gas = ct.Solution('h2-co-mechanism.cti')
gas.TPX = T_0, P_0, X_0

reac = ct.IdealGasReactor(gas)
env = ct.Reservoir(ct.Solution('air.xml'))

time = dp.volume_history.time
volume = dp.volume_history.volume
ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))

netw = ct.ReactorNet([reac])
while netw.time < 0.05:
    netw.step()

\end{minted}

@@

The \ck{} file used in this example can be found online.\todo{add url}{}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol.
Examining the file, we first find the section that specifies the information about the \ck{} file itself:
%
\begin{minted}{yaml}
file-author:
  name: Morgan Mayer
  ORCID: 0000-0001-7137-5721
file-version: (0,1)
chemked-version: 0.0.1
screw-you-atom: 0
\end{minted}
%
Then the reference information:
%
\begin{minted}{yaml}
reference:
  doi: 10.1016/j.combustflame.2011.08.014
  authors:
    - name: Ivo Stranic
    - name: Deanna P. Chase
    - name: Joseph T. Harmon
    - name: Sheng Yang
    - name: David F. Davidson
    - name: Ronald K. Hanson
  journal: Combustion and Flame
  year: 2011
  volume: 159
  pages: 516-527
  detail:
experiment-type: ignition delay
apparatus:
  kind: shock tube
  institution: High Temperature Gasdynamics Laboratory, Stanford University
  facility: stainless steel shock tube
dammit-atom: 0
\end{minted}
%
In this \ck{} file, multiple data points are specified with some common
conditions, including a common mixture composition and common defintion of
ignition delay. Therefore, a \yaml|common-properties| section is specified.
%
\begin{minted}{yaml}
common-properties:
  composition: &comp
    - species-name: t-butanol
      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
      mole-fraction: 0.003333333
    - species-name: O2
      InChI:  1S/O2/c1-2
      mole-fraction: 0.04
    - species-name: Ar
      InChI:  1S/Ar
      mole-fraction: 0.956666667
  ignition-type:  &ign
    target: OH* concentration
    type: 1/2 max
atom-is-bad-at-syntax-highlighting: 0
\end{minted}
%
This block utilizes the ability of YAML files to define an anchor using the
\yaml|&| symbol and refer to that section later using the \yaml|*| symbol, as
can be seen in the definition of the \yaml|datapoints| section:
%
\begin{minted}{yaml}
datapoints:
  - temperature: 1459 kelvin
    ignition-delay: 347 us
    pressure: 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1389 kelvin
    ignition-delay: 756 us
    pressure: 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1497 kelvin
    ignition-delay: 212 us
    pressure: 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1562 kelvin
    ignition-delay: 105 us
    pressure: 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{minted}

In this example, we would like to run constant volume simulations at each of the
pressure and temperature conditions in the \yaml|datapoints| list.

\begin{minted}{python}
import cantera as ct
from pyked import ChemKED

@@

X_0 = ck.datapoints[0].get_cantera_mole_fraction()

for dp in ck.datapoints:
    T_0 = dp.temperature.to('K').magnitude
    P_0 = dp.pressure.to('Pa').magnitude
    gas.TPX = T_0, P_0, X_0
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T_0 + 400:
        netw.step()

    print('The ignition delay for T_0={}, P_0={} is: {} seconds'.format(
        T_0, P_0, netw.time
    ))
\end{minted}
%====================================================================
\section{Conclusions and future work}
%
\todo[inline]{both to do}
